*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-15 Code/Manuscript_O2.Rmd
Mireille, Aurora, introduction merge

*Prochlorococcus marinus*, a genus of the phylum Cyanobacteria, is the smallest known photosynthetic prokaryote, with cell diameters ranging from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992]. Despite a small cell size, *P. marinus* contribute 13 to 48% of net primary production in oligotrophic oceans, corresponding to about 30% of  global oxygen production [@partenskyProchlorococcusMarinePhotosynthetic1999].

*Prochlorococcus marinus* comprises many strains, organized into clades,
which in turn are grouped into ecotypes defined largely as High-Light (HL) or Low-Light (LL) adapted,  and aligned phylogenetically [@moorePhysiologyMolecularPhylogeny1998]. The biomass distribution of *P. marinus* is constrained to between latitudes of 40°N to 40°S in open ocean waters [@partenskyProchlorococcusMarinePhotosynthetic1999] with different ecotypes inhabiting distinct ecological niches. High-Light ecotypes are typically observed as the dominant picophytoplankton in near-surface oligotrophic waters that are, characterized by low nutrient concentrations, and high light levels. The monophyletic high light ecotype is subdivided into clades based on iron adaptation, whereby HLIII/IV are low iron adapted, whereas HLI is a high iron, low temperature adapted clade and HLII is a high iron, high temperature adapted clade [@kentParallelPhylogeographyProchlorococcus2019; @johnsonNichePartitioningProchlorococcus2006; @zinserInfluenceLightTemperature2007].  Low-Light ecotypes thrive in deeper ocean waters, extending beyond 200 m in depth [@partenskyProchlorococcusMarinePhotosynthetic1999]. In these deeper regions, only 1% of the surface irradiance penetrates, primarily in the blue (450 nm) to green (520 nm) spectral waveband ranges, due to light and spectral attenuation [@holtropVibrationalModesWater2021]. Additionally, some *P. marinus* strains in clade LLV and LLVI thrive in  Oxygen Minimum Zones (OMZ) of the subtropical Atlantic and Pacific Oceans, where dissolved oxygen concentrations [O~2~] can be less than 20 µM [@ulloaCyanobacteriumProchlorococcusHas2021]. Currently there are at least 12 clades but only 5 have been cultivated; HLI, HLII, LLI, LLII/III and LIV [@billerProchlorococcusStructureFunction2015].

Our changing climate, characterized by warming ocean waters, is rapidly altering conditions for these highly specialized strains of marine picophytoplankton. Predictions indicate a net global increase of *P. marinus* cell abundances of 29% and poleward latitudinal shifts of at least 10° in marine phytoplankton niches by the end of this century [@flombaumPresentFutureGlobal2013] in response to warming waters, with cell abundance increases of approximately 50% in the more poleward regions. As the solar incidence angles are lower and photoperiod seasonality is more pronounced at higher latitudes, this shift will carry these phytoplankters into new combinations of photoperiod, seasonal regimes and changing spectral regimes [@bartonAnthropogenicClimateChange2016]. A 10° poleward shift of *P. marinus*, coupled with a dramatic increase in cell abundance between 40°N and 50°N,  could have a large impact on ocean ecosystems and global biogeochemical cycles [@flombaumPresentFutureGlobal2013], and thus will undoubtedly impact the ecosystem balance off our Canadian waters.

Global warming is also rapidly changing the chemistry of our oceans.  By the end of this century, surface ocean pH is projected to decline by 0.1 to 0.4  due to  projected increases in carbon dioxide concentrations [@garcia-sotoOverviewOceanClimate2021]. Moreover, substantial changes in the global water cycle, leading to extensive changes in worldwide precipitation patterns, are affecting ocean salinity levels on a global scale, and  ice melts due to rising temperatures are impacting salinity levels in the Arctic and Northwest Atlantic oceans [@leeRecognizingSalinityThreats2022]. There is evidence that increasing sea temperatures are also causing a decrease in [O~2~] across our global oceans [@matearLongtermChangesDissolved2003], with significant declines observed toward the poles in both hemispheres [@helmObservedDecreasesOxygen2011]. Warmer ocean waters decrease oxygen solubility at the surface which in turn reduces the amount of oxygen that can be mixed downwards by ocean currents [@garcia-sotoOverviewOceanClimate2021]. Furthermore, models predict that the OMZ in the Pacific and Indian Oceans are expanding [@garcia-sotoOverviewOceanClimate2021; @buseckeDivergingFatesPacific2022]. In light of these expanding OMZ, a recent study by @buseckeDivergingFatesPacific2022 highlights an intriguing finding; the core of the OMZ, where the oxygen levels are lowest, is projected to contract.  Given the predicted decrease in overall oceanic [O~2~],   projected expansion of OMZs, and possibly decline in the OMZ core volumes, we set forth to examine the influence of [O~2~] on *P. marinus* growth as this area of research has been relatively understudied.

To determine the viability of *P. marinus* strains in changing ecosystems, we analyzed the growth and physiological responses of representative strains from three clades to varying [O~2~], light levels, spectral waveband ranges, and photoperiods,  to simulate a matrix of eco-physiological conditions representative of current and hypothetical future ocean zones. The following *P. marinus* clades were examined: *P. marinus* MED4, classified as a clade HLI (Figure \@ref(fig:PhylogenyTreeYan2018edit)), was isolated near the ocean surface (5m depth) of the Mediterranean Sea where [O~2~] are near saturation, light levels are high and spectral bias from full solar irradiance is minimal. *P. marinus* SS120, a clade LLIII (Figure \@ref(fig:PhylogenyTreeYan2018edit)), was isolated from the Sargasso Sea at a depth of 120 m, while *P. marinus* MIT9313, a clade LLIV (Figure \@ref(fig:PhylogenyTreeYan2018edit)), was isolated from the North Atlantic Gulf Stream at a depth of 135 m. At these depths, light attenuation and spectral shifts occur, resulting in low blue and green light levels, while [O~2~] varies from near-surface saturation levels to decreased concentrations, but it does not necessarily decrease with depth [@billheimerOxygenSeasonalityUtilization2021]. In chapter 2, we define the interactions between oxygen, light level, light colour and photoperiod that help define their complex niches.

In chapter 3, to better decipher the physiological processes driving the observed growth responses to a complex matrix of spectral light regimes and [O~2~], we express growth rates in terms of cumulative diel Photosynthetically Usable Radiation (PUR). Photosynthetic organisms rely on absorbing light energy within the visible spectrum, 350 to 700 nm, for photosynthesis. This range of light, known as Photosynthetic Active Radiation (PAR) [@morelAvailableUsableStored1978], is essential for the photosynthetic process. PUR represents the fraction of PAR that is absorbed by the pigments of photosynthetic organisms [@morelAvailableUsableStored1978], thus we can compare light levels across different spectral wavebands that are absorbed differently by cells. Moreover, expressing growth rates in terms of cumulative diel absorbed PUR has the potential to simplify different photoperiods, spectral bands, and PAR levels into a common parameter, making growth response comparisons across different oxygen levels more attainable and may allow us to determine whether growth responses are primarily influenced by cumulative diel absorbed PUR, specific photoperiods or spectral bands.

This study focuses on evaluating how growth of *Prochlorococcus* strains from different clades responds to varying [O~2~], spectral wavebands, light levels and photoperiods.

### Chapter 2 PAR Intro
*Prochlorococcus marinus* ecotypes are characterized by their distinct
tolerance for High-Light (HL) or Low-Light (LL) levels
[@moorePhysiologyMolecularPhylogeny1998]. *Prochlorococcus marinus* is
classified into at least 12 major clades
[@billerProchlorococcusStructureFunction2015], that respectively inhabit
a range of ecological niches across vertical ocean water columns
[@chisholmNovelFreelivingProchlorophyte1988;
@veldhuisVerticalDistributionPigment1990;
@partenskyProchlorococcusMarinePhotosynthetic1999].

Previous studies have observed that the diversification of *P. marinus*
niches extends into regions considered unfavorable for photosynthesis to
occur. @goerickeNovelNicheProchlorococcus2000 first reported the
existence of *P. marinus* below the oxyclines of the Oxygen Minimum
Zones (OMZ) of the Arabian Sea and the Eastern Tropical North Pacific.
In an OMZ, oxygen concentrations [O~2~] can be less than 20 µM, and in
some regions [O~2~] is undetectable
[@garcia-robledoCrypticOxygenCycling2017a]. Subsequent research has
revealed that *P. marinus* LL ecotypes dominate the OMZ
[@lavinNovelLineagesProchlorococcus2010;
@ulloaCyanobacteriumProchlorococcusHas2021] and are believed to
contribute to approximately 90% of the *P. marinus* diversity in the
eastern tropical Pacific Ocean OMZ
[@partenskyProchlorococcusAdvantagesLimits2010].

Although *P. marinus* ecotypes are defined by their ecological niches,
there is evidence to suggest that ecotypes can exist in environments
beyond their optimal habitats. HL ecotypes are known to
inhabit depths where LL ecotypes also inhabit
[@westNichePartitioningProchlorococcusPopulationsStratified1999;
@zinserInfluenceLightTemperature2007;
@delmontLinkingPangenomesMetagenomes2018] but at light levels still
within the range of confirmed growth irradiance limits
[@moorePhotophysiologyMarineCyanobacterium1999]. Studies by
@lavinNovelLineagesProchlorococcus2010 and
@mattisonPhotosyntheticStrategiesWild2020 show evidence that LL ecotypes
can occupy regions in the OMZ at depths above 40 m. Since light levels
were not reported at every depth of these studies, it is presumed that
at a depth of 40 m, closer to the surface, light levels are higher than
what LL ecotypes can tolerate. Furthermore,
@mattisonPhotosyntheticStrategiesWild2020 presented evidence of proteins
derived from HL ecotypes of *P. marinus* in the OMZ at depths up to 200
meters with O~2~ of 15 µM. However, the extent to which [O~2~] defines
the niches occupied by different *P. marinus* ecotypes, as compared to
potentially covarying environmental variables like photoperiod, light
spectrum, and light level, is poorly described.

Photosynthetic organisms rely on the absorption of light energy within
the visible spectral range of 350 to 700 nm to drive the photosynthetic
process. This spectral range of light, known as Photosynthetic Active
Radiation (PAR) [@morelAvailableUsableStored1978], is essential for
photosynthesis. However, as sunlight enters the marine environment, it
undergoes attenuation, resulting in a decline of specific wavebands of
light through the water column from the surface. This attenuation
primarily affects the longer wavelengths in the red region of the
spectrum (600 to 700 nm), followed by wavelengths in the green region
(500 to 600 nm) and, finally, wavelengths in the blue region (400 to 500
nm) [@holtropVibrationalModesWater2021]. Thus, as sunlight penetrates the
marine environment, the light level diminishes and the residual spectral
wavebands shift with depth, where only blue spectral wavelengths
dominate. The region of the water column with enough light to support
photosynthesis is known as the photic zone, defined as the depth where
light levels remain above 20 µmol photons m^-2^ s^-1^, 1% of the surface
solar incident radiation, of 2000 µmol photons m^-2^ s^-1^.

In growth studies of *P. marinus*, it is common to use a broad spectrum
light source to approximate the solar spectral wavelengths found in the
uppermost layers of the ocean. The actinic photon flux is adjusted to
simulate the specific light levels appropriate for the ecotype of
interest. However, this approach falls short when attempting to
accurately reproduce the ambient spectral light quality of deep ocean
habitats, where only blue spectral wavelengths prevail
[@holtropVibrationalModesWater2021]. *Prochlorococcus marinus* contain a
unique light-harvesting complex composed primarily of divinyl
chlorophyll a and b [@chisholmProchlorococcusMarinusNov1992;
@goerickePigmentsProchlorococcusMarinus1992], the only photosynthetic
organism to possess these chlorophyll derivatives. The absorption maxima
of 442 nm for divinyl chlorophyll a and 478 nm for divinyl chlorophyll b
[@goerickePigmentsProchlorococcusMarinus1992] allow *P. marinus* to
efficiently harvest the light energy emitted by blue light in the 400 nm
to 500 nm range [@morelAvailableUsableStored1978].

At most latitudes, the duration of the photoperiod (day length)
undergoes seasonal variations. However, at the equator, the photoperiod
remains relatively constant at the ocean surface, approximately 12 hours
(h) of daylight and 12 h of darkness throughout the entire year. As *P.
marinus* potentially expands its temperature-permissive niches poleward
into temperate regions due to climate change
[@flombaumPresentFutureGlobal2013;
@bartonAnthropogenicClimateChange2016], it will encounter more
pronounced seasonal variations in photoperiod regimes. Photoperiods are
longer at temperate regions during the summer, while in spring, fall and
winter, photoperiods are shorter. Additionally, when considering the
influence of light attenuation with depth, temperate regions exhibit
more combinations of photoperiod compared to equatorial regions. The
effects of photoperiod on the ecophysiology of phytoplankton is
underscored by several studies. @liDiatomGrowthResponses2017
demonstrated complex interactions of photoperiod and growth rate of
diatoms. Similarly, @prezelinDielPeriodicityPhytoplankton1992 expressed
the importance of considering photoperiod on *in situ* rates of primary
production, as the diel light to dark ratio regulates metabolic
processes throughout the day. Furthermore,
@vaulotGrowthProchlorococcusPhotosynthetic1995 showed that
*Prochlorococcus* undergoes DNA replication in the afternoon, while cell
division occurs at night. Further research is needed to investigate the
impact of different photoperiods on the growth of *Prochlorococcus*, as
this aspect remains inadequately studied.

@@

(fig:SunDepth) shows light levels and spectral waveband
attenuation at ocean surface to ocean depths of 1000 m, as well as the
hours of light above the minimum photic zone of 20 µmol photons m^-2^
s^-1^, representing the photoperiod. This figure is modeled using the
solar incident angle at 45°N, a hypothetical future niche, at summer
solstice and illustrates the light parameters that define the
photoregimes of the niches of *P. marinus* that we studied. Figure (fig:CardinalSeasons) illustrates light attenuation through ocean waters at 45°N at winter solstice, equinox and summer solstice.

### Chapter 3 Intro (PUR)

*Prochlorococcus marinus* ecotypes are characterized by their distinct
tolerance for High-Light (HL) or Low-Light (LL) levels. The photosynthetic strategies between *Prochlorococcus marinus* ecotypes must differ in order to exploit the very diverse photo regimes in their respective niches and may be explained by their unique cellular compositions. 
*Prochlorococcus marinus* MED4 has the smallest genome size of the 3 ecotypes we studied, only 1.66 Mbp, while *P. marinus* MIT9313 has the largest genome size of 2.41 Mbp [@rocapGenomeDivergenceTwo2003]. *Prochlorococcus marinus* SS120 genome size is 1.75 Mbp [@dufresneGenomeSequenceCyanobacterium2003]. Due to the small genome size and small cell diameters, ranging from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992], *Prochlorococcus* are the smallest known photosynthetic prokaryote. The simplistic cellular structure and small size of *P. marinus* minimizes the pigment package effect or intracellular self-shading, due to the high surface area to volume ratio, contributing to very efficient optical absorption capabilities [@morelProchlorococcusSynechococcusComparative1993].

Photosynthetic organisms depend on the absorption of light energy in the visible spectrum, 350 to 700 nm, to carry out photosynthesis. This range of light, referred to as Photosynthetic Active Radiation (PAR) [@morelAvailableUsableStored1978], is crucial for the photosynthetic process. The Photosynthetic Usable Radiation (PUR) is defined as the fraction of PAR absorbed by the pigments of a given photosynthetic taxon [@morelAvailableUsableStored1978]. PUR, therefore, takes into account the pigment composition of photosynthetic organisms and the specific spectral wavebands that these pigments absorb. *Prochlorococcus marinus* MED4, a clade HLI ecotype, dominate the near ocean surface where spectral attenuation from full solar irradiance is minimal and light levels are high [@partenskyProchlorococcusMarinePhotosynthetic1999]. *Prochlorococcus marinus* SS120, a clade LLIII ecotype, and *P. marinus* MIT9313, a clade LLIV ecotype, dominate the deep ocean [@partenskyProchlorococcusMarinePhotosynthetic1999] where extreme light attenuation and spectral shifts occur, resulting in the availability of very low light levels in the blue (450 nm) to green (520 nm) spectral range [@holtropVibrationalModesWater2021]. Given the different spectral light regimes of the different ecotype niches, understanding how the quantity of available light levels, PAR, converts to what these ecotypes can use, PUR, can give some insight into how they are able to exploit light energy.

The photosynthesis process can be divided into two distinct reactions: the
light-dependent and the light-independent reactions. *Prochlorococcus marinus* does not contain
chloroplasts [@chisholmNovelFreelivingProchlorophyte1988], and the
light-dependent reactions occur in thylakoid membranes arranged concentrically around the periphery of the cell. Light harvesting complexes spanning a thylakoid membrane capture or 'harvest' light
energy. The light harvesting complexes in *P. marinus* consist of Prochlorophyte Chlorophyll Binding
proteins (Pcb) [@larocheIndependentEvolutionProchlorophyte1996]. Once photons are absorbed by light harvesting pigments, the energy is transferred to the reaction center of Photosystem II
(PSII) where water is split, catalyzed by the manganese cluster of the
oxygen evolving complex. Here, 2 water molecules are oxidized to produce 1 oxygen
(O~2~) molecule, releasing 4 protons into the thylakoid lumen, which contribute to the transmembrane proton
potential,  and 4 electrons. These electrons are passed to Photosystem I
(PSI) through a series of reduction and oxidation reactions in the
electron transport chain in the thylakoid membrane, and in the process, deliver more protons to the
proton pool within the lumen side of the thylakoid membrane. Photosystem I
(PSI) also harvest light energy but does not split water, rather the
reaction center uses excitation energy to excite electrons received from PSII
and transfer them by the reduction and oxidation of iron-sulphur
clusters to reduce NADP^+^ to NADPH. The Adenosine Triphosphate (ATP) synthase complex produces ATP in the stroma, driven by protons passing from the lumen out to the stroma side of the thylakoid
membrane. The light-independent reaction involves the RuBisCO enzyme
which uses the NADPH and ATP generated in the light-dependent reaction
to convert dissolved carbon dioxide (CO~2~) gas into carbohydrates.

*Prochlorococcus marinus* is differentiated from other cyanobacteria as a distinct lineage, primarily due to the absence of phycobilisomes [@chisholmProchlorococcusMarinusNov1992]. Instead, *Prochlorococcus*  utilizes intra-membrane Pcb proteins bound to divinylchlorophyll a (Chl a2), and α-carotene, as the major light harvesting complex [@goerickePigmentsProchlorococcusMarinus1992]; [@larocheIndependentEvolutionProchlorophyte1996].  Not all *Prochlorococcus* encode the same number of Pcb proteins. The representative of HLI clade, *P. marinus* MED4 has 1 gene encoding pcbA protein [@garczarekMultiplicationAntennaGenes2000] and the LLIV clade, *P. marinus* MIT9313, has 2 genes encoding pcbB and pcbG proteins [@rocapGenomeDivergenceTwo2003]. The LLIII clade, *P. marinus* SS120, however, has 7 genes encoding pcbA, pcbB, pcbC, pcbE, pcbF, pcbG and pcbH [@garczarekMultiplicationAntennaGenes2000]. According to @garczarekMultiplicationAntennaGenes2000, the expression of some of the pcb genes might be specifically induced under low light conditions, after noting the thickening of the antenna band of the thylakoid membrane when SS120 was switched to low light after high light exposure. Another difference within the cell structure among *P. marinus* ecotypes is the 2 to 4 thylakoids that run parallel to the cell membrane. For SS120, the thylakoids are closed around the periphery of the cell; however, in MED4 and MIT9313, there is a gap resulting in a horseshoe shape for the thylakoids [@partenskyProchlorococcusMarinePhotosynthetic1999].  Despite the deviation in the light harvesting complex from cyanobacteria, the genes encoding the rest of photosynthetic apparatus is highly conserved in *Prochlorococcus* [@kettlerPatternsImplicationsGene2007], though the photosynthetic efficiency may vary between ecotypes [@hessPhotosyntheticApparatusProchlorococcus2001; @moorePhotophysiologyMarineCyanobacterium1999].

>>>>

*Prochlorococcus marinus*, a genus of the phylum Cyanobacteria, is the smallest known photosynthetic prokaryote, with cell diameters ranging from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992]. Despite small cell size, *P. marinus* contribute 13 to 48% of net primary production in oligotrophic oceans, corresponding to about 30% of  global oxygen production [@partenskyProchlorococcusMarinePhotosynthetic1999].  *P. marinus* growth is currently constrained to between latitudes of 40°N to 40°S in open ocean waters, from surface to 300 m  depth, thus spanning 3 orders of magnitudes of light (Partensky et al., 1999) (Chisholm et al., 1992)  [@partenskyProchlorococcusMarinePhotosynthetic1999].

*Prochlorococcus marinus* is functionally differentiated from other cyanobacteria primarily by the absence of phycobilisomes [@chisholmProchlorococcusMarinusNov1992]. Instead, *Prochlorococcus*  utilizes intra-membrane Pcb proteins bound to divinylchlorophyll a (Chl a2), and α-carotene, as the major light harvesting complex [@goerickePigmentsProchlorococcusMarinus1992]; [@larocheIndependentEvolutionProchlorophyte1996], with diversity in the number of genes encoding Pcb, expression of Pcb under changing light [@garczarekMultiplicationAntennaGenes2000]  [@rocapGenomeDivergenceTwo2003] [@garczarekMultiplicationAntennaGenes2000], and in thylakoid organization [@partenskyProchlorococcusMarinePhotosynthetic1999].  the genes encoding most of the rest of the photosynthetic apparatus is highly conserved in *Prochlorococcus* [@kettlerPatternsImplicationsGene2007], though the photosynthetic efficiency may vary among clades [@hessPhotosyntheticApparatusProchlorococcus2001; @moorePhotophysiologyMarineCyanobacterium1999].

*Prochlorococcus marinus* comprises many strains, organized into clades, defined by 16S-23S intergenic transcribed ribosomal sequence signatures (Rocap et al., 2003). The various clades inhabit distinct ecological niches [@moorePhysiologyMolecularPhylogeny1998], originally defined by Low Light or High Light.  Of at least 12 known *Prochlorococcus* clades only 5 to date have cultured representatives; HLI, HLII, LLI, LLII/III and LIV [@billerProchlorococcusStructureFunction2015].

XXXXMIREILLE Check strain descriptions for accuracyXXXX
Low-Light clades thrive in deeper ocean waters, extending beyond 200 m in depth [@partenskyProchlorococcusMarinePhotosynthetic1999], where only ~1% of the surface irradiance penetrates, primarily in the blue (450 nm) to green (520 nm) spectral ranges [@holtropVibrationalModesWater2021]. Clade LLIV, including cultured strain MIT9313, falls near the base of the *Prochlorococcus* radiation, and is characterized by preference for low light, at depths from 120 m and 200 m (Partensky et al., 1999). LLIV members are, as yet, the only cultured strains found in Oxygen Minimum Zones (OMZ).  Some as yet uncultured *P. marinus* strains in clades LLV and LLVI thrive in OMZ of the subtropical Atlantic and Pacific Oceans, where dissolved oxygen concentrations [O~2~] can be less than 20 µM [@ulloaCyanobacteriumProchlorococcusHas2021] (Lavin et al., 2010) ADDITIONAL REFS XXXX.   (Johnson et al., 1999). @goerickeNovelNicheProchlorococcus2000
[@garcia-robledoCrypticOxygenCycling2017a]. *P. marinus* LL ecotypes may indeed dominate the phytoplankton wihthin OMZ
[@lavinNovelLineagesProchlorococcus2010;
@ulloaCyanobacteriumProchlorococcusHas2021]
[@partenskyProchlorococcusAdvantagesLimits2010].

Clades LLII and LLIII are often grouped together as the second oldest phylogenetic lineages in the *Prochlorococcus* radiation, and share an affinity for low light; LLIII cultured strain SS120, was isolated from 135 m depth in the Sargasso Sea.

Clade LLI  includes cultured strains NATL1A and NATL2A, which prefer moderate irradiances found between 30 and 100 m depth.

HL clades are the most recently evolved lineages, with reduced genome sizes in comparison to LL ecotypes. High-Light clades are  typically dominant picophytoplankton in near-surface, oligotrophic waters, characterized by high light levels. Clade HLI, including cultured strain MED4,is adapted to higher iron, and lower temperatures, and originated from the Mediterranean Sea, just 5 m below the surface (Partensky et al., 1999).  Clade HLII adapted to higher iron, and higher temperatures, is the most abundant group in the North Atlantic and North Pacific Oceans, often constituting over 90% of the total population (Partensky et al., 1999),  and are most numerous around 50 m depth (Partensky et al., 1999). Clade HLIII/IV is adapted to lower iron [@kentParallelPhylogeographyProchlorococcus2019; @johnsonNichePartitioningProchlorococcus2006; @zinserInfluenceLightTemperature2007].

*P. marinus* clades are found  in environments beyond their optimal habitats. HL clades inhabit depths overlapping with LL ecotypes
[@westNichePartitioningProchlorococcusPopulationsStratified1999;
@zinserInfluenceLightTemperature2007;
@delmontLinkingPangenomesMetagenomes2018] 
[@moorePhotophysiologyMarineCyanobacterium1999]. 
@lavinNovelLineagesProchlorococcus2010 and
@mattisonPhotosyntheticStrategiesWild2020 show that LL ecotypes
can occupy regions in the OMZ at depths above 40 m, reaching ambient light levels above what LL clades were thought to tolerate.

*Prochlorococcus* remain challenging to culture, as their reduced genomes – the smallest of any known oxyphototroph – render them partially dependent upon mutualistic heterotrophic bacteria to  detoxify reactive oxygen species (Moore et al., 2007; Morris et al., 2008). MED4, SS120 and MIT9313 have been successfully cultured in laboratories, and used to show that ecotypic classifications correspond to biochemical differences among strains. CITATONS OUR LAB and OTHERS.

Our changing climate, characterized by warming ocean waters, is rapidly altering conditions for these highly specialized strains of marine picophytoplankton. Predictions indicate a net global increase of *P. marinus* cell abundances of 29%, along with  poleward latitudinal shifts of at least 10° in marine phytoplankton niches by the end of this century [@flombaumPresentFutureGlobal2013] in response to warming waters, with increases in *Prochlorococcus* of approximately 50% in the more poleward regions. As the solar incidence angles are lower and photoperiod seasonality is more pronounced at higher latitudes, this shift will carry these phytoplankters into new combinations of photoperiod, seasonal regimes and changing spectral regimes [@bartonAnthropogenicClimateChange2016]. A 10° poleward shift of *P. marinus*, coupled with a dramatic increase in cell abundance between 40°N and 50°N,  could have a large impact on ocean ecosystems and global biogeochemical cycles [@flombaumPresentFutureGlobal2013], and thus will undoubtedly impact the ecosystem balance off our Canadian waters.

Global warming is also rapidly changing the chemistry of our oceans.  By the end of this century, surface ocean pH is projected to decline by 0.1 to 0.4  due to  projected increases in carbon dioxide concentrations [@garcia-sotoOverviewOceanClimate2021]. Moreover, substantial changes in the global water cycle, leading to extensive changes in worldwide precipitation patterns, are affecting ocean salinity levels on a global scale, and  ice melts due to rising temperatures are impacting salinity levels in the Arctic and Northwest Atlantic oceans [@leeRecognizingSalinityThreats2022]. Increasing sea temperatures are also causing a decrease in [O~2~] across our global oceans [@matearLongtermChangesDissolved2003], with significant declines in [O~2~] observed toward the poles in both hemispheres [@helmObservedDecreasesOxygen2011]. Warmer ocean waters increase stratification, and decrease oxygen solubility at the surface, which in turn decreases oxygen mixing downwards by ocean currents [@garcia-sotoOverviewOceanClimate2021]. Models predict that the OMZ in the Pacific and Indian Oceans are expanding [@garcia-sotoOverviewOceanClimate2021; @buseckeDivergingFatesPacific2022]. In light of these expanding OMZ, a recent study by @buseckeDivergingFatesPacific2022 highlights an intriguing finding; the cores of the OMZ, where the oxygen levels are lowest, is projected to contract.  Given the predicted decrease in overall oceanic [O~2~],projected expansion of OMZs, and possible decline in the OMZ core volumes, we set forth to examine the influence of [O~2~] on *P. marinus* growth, as this area of research has been relatively understudied.

Near the equator, photoperiod remains relatively constant at the ocean surface, approximately 12 hours (h) of daylight and 12 h of darkness throughout the entire year. The effective length of the photoperiod does, however attentuate with depth as dawn and dusk light drops below the level needed for biological processes. As *P. marinus* potentially expands its temperature-permissive niches poleward into temperate regions 
[@flombaumPresentFutureGlobal2013; @bartonAnthropogenicClimateChange2016], it will encounter more
pronounced seasonal variations in photoperiod regimes at surface and at depth. With the influence of light attenuation with depth, temperate regions exhibit more combinations of photoperiod, compared to equatorial regions, with potentially complex effects upon viable growth niches @liDiatomGrowthResponses2017 @prezelinDielPeriodicityPhytoplankton1992.  @vaulotGrowthProchlorococcusPhotosynthetic1995 showed that *Prochlorococcus* replication of DNA occurs in the afternoon, while cell division occurs at night.

We therefore analyzed the growth and physiological responses of representative strains from three clades under a matrix of [O~2~], light levels, spectral waveband ranges, and photoperiods,  to approximate eco-physiological conditions representative of current and hypothetical future ocean zones. *P. marinus* MED4, a clade HLI strain (Figure \@ref(fig:PhylogenyTreeYan2018edit)), was isolated near the ocean surface (5m depth) of the Mediterranean Sea where [O~2~] is near saturation, light levels are high and spectral bias from full solar irradiance is minimal. *P. marinus* SS120, a clade LLIII (Figure \@ref(fig:PhylogenyTreeYan2018edit)), was isolated from the Sargasso Sea at a depth of 120 m, while *P. marinus* MIT9313, a clade LLIV (Figure \@ref(fig:PhylogenyTreeYan2018edit)), was isolated from the North Atlantic Gulf Stream at a depth of 135 m. At these depths, light attenuation and spectral shifts occur, resulting in low blue and green light levels, while [O~2~] varies from near-surface saturation levels to decreased concentrations, but does not necessarily decrease systematically with depth [@billheimerOxygenSeasonalityUtilization2021].

We then analyzed growth rates in terms of cumulative diel Photosynthetically Usable Radiation (PUR). Photosynthetic organisms rely on absorbing light energy within the PAR range, 350 to 700 nm, for photosynthesis [@morelAvailableUsableStored1978]. PUR represents the fraction of PAR that is absorbed by the pigments of photosynthetic organisms [@morelAvailableUsableStored1978]. PUR, therefore, takes into account the pigment composition of photosynthetic organisms and the specific spectral wavebands these pigments absorb.  *Prochlorococcus marinus* light-harvesting complexes, [@chisholmProchlorococcusMarinusNov1992;
@goerickePigmentsProchlorococcusMarinus1992], show an absorption maxima
of 442 nm for divinyl chlorophyll a and 478 nm for divinyl chlorophyll b
[@goerickePigmentsProchlorococcusMarinus1992] allowing *P. marinus* to
efficiently harvestblue light in the 400 nm
to 500 nm range [@morelAvailableUsableStored1978] prevailing in deep ocean
habitats, where only blue spectral wavelengths prevail [@holtropVibrationalModesWater2021].  In *Prochlorococcus* small cell diameters, from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992], and simple cell structures, minimize the complication of pigment package effect or intracellular self-shading [@morelProchlorococcusSynechococcusComparative1993]. Given the different spectral light regimes typical of the niches of different ecotypes,  expressing growth rates in terms of cumulative diel absorbed PUR might simplify different photoperiods, spectral bands, and PAR levels into a common parameter, making growth response comparisons across strains and different oxygen levels more accessible.  We also aimed to detect whether growth responses are driven simply by cumulative diel absorbed PUR, or whether specific photoperiods, spectral bands or PAR levels have independent, albeit interacting, effects on growth.

@@

Under full atmospheric [O~2~], LL clades of *Prochlorococcus* are restricted to growth under low light, in part because they suffer photoinhibition of Photosystem II (PSII) through several paths, including direct absorbance of UV or blue light, in parallel with generation of Reactive Oxygen Species (ROS) if the electron flow is slowed, [@aroPhotoinhibitionPhotosystemII1993], producing  damaging singlet oxygen (^1^O~2~). [@aroPhotoinhibitionPhotosystemII1993; @soitamoPhotoinhibitionMarinePicocyanobacteria2017] [@murphyPhotoinactivationPhotosystemII2017][@hakalaEvidenceRoleOxygenevolving2005].
Repair of inhibited PSII process relies on the removal of damaged PsbA (Mann et al., 2000; Adam et al., 2005), followed by reassembly with newly synthesized PsbA [@nixonRecentAdvancesUnderstanding2010]. 
Thus, degradation of PsbA is a rate-limiting step in recovery from photoinhibition (Kanervo et al., 1993), mediated largely by a heterohexamer of (FtsH12)3, a membrane-bound (Sacharz et al., 2015)(Zak et al., 2001) metalloprotease  (Chiba et al., 2002) (Yoshioka-Nishimura and Yamamoto, 2014).  (Adam et al., 2005; Boehm et al., 2012) (Nixon et al. (2010)) (Komenda et al., 2007). (Pisareva et al., 2007; Sacharz et al., 2015),  (Mann et al., 2000), (AMANDA REF from ALGATECH).

*Prochlorococcus* genomes encode 4 FtsH proteins, henceforth referred to as FtsH1-4, homologous to the characterized FtsH isoforms of the model cyanobacterium *Synechocystis* FtsH, with presumably parallel functions (Table 1).  Upon a shift to higher light HLIII MED4 upregulated expression of FtsH1 and FtsH2 (Bonisteel et al. (2018)), homologs to the *Synechocystis* slr0228 and slr1604, implicated in PSII repair. In contrast representative LLIII strain MIT9313 showed less overall expression of the FtsH proteases, and thus has fewer FtsH serving each photosystem.  Furthermore MIT9313 expressed primarily FtsH3, homologous to Sll1463 involved in PSI biogenesis, and FtsH expression did not increase in response to light stress. Through adaptation to steady low light, LLIII **Prochlorococcus* instead allocates resources processes other than dynamic regulation of PSII repair.

```{r FtsHtable, table.cap = FtsHtable_cap, echo = FALSE}

FtsHtable_cap <- glue("FtsH protease homologs in *Prochlorococcus marinus* and the model cyanobacterium *Synechocystis* sp. PCC6803. Protein homology determined by multiple sequence alignment with MUSCLE followed by construction of maximum likelihood phylogenetic tree using 1000 bootstrap replicates in MEGAX.",

                       .sep = " ")
FtsHTable <- tibble(Organism = c("Prochlorococcus marinus", "Synechocystis sp. PCC6803", "Function"),
                    FtsH1 = c("FtsH1", "SlrO22", "PSII Repair"),
                    FtsH2 = c("FtsH2", "Slr1604", "PSII Repair"),
                    FtsH3 = c("FtsH3", "Slr1463", "PSI biogenesis"),
                    FtsH4 = c("FtsH4", "Slr1390", "Cell viability")
                    )

FtsHTable
```
